---
output: pdf_document
---

```{r, message = FALSE, echo = FALSE}
library(tidyverse)
library(ggdist)
library(distributional)
```

# Chapter 4 R Practice

## Uniform distribution: Suppose $X \sim U(0,5)$.

1.  Sketch the pdf and shade the region $X < 2$

\
\
\


2.  Write out the integral and the R code required to calculate $P(X < 2)$. Calculate it using your preferred method.
    \
    \
    \
\
\
\

3.  Write out two different lines of code you could use to find
    $P(X > 2)$ in R. *Hint: one should make use of the `lower.tail`
    argument.* Execute both lines of code in R to verify that they give
    the value you expect.\
    \
    \
    \
    \
    \

4.  Suppose we want to find $P(2 < X < 3)$. Sketch a graph with this
    region shaded.

\
\
\
\
\


5.  Using your visual as a guide, what would you need to subtract from
    $P(X < 3)$ in order to obtain $P(2 < X < 3)$?

\
\
\

6.  Using your answer to part 5 as a guide, write out R code for
    computing $P(2 < X < 3)$. Execute your code in R and report the
    answer.

## Normal distribution

7. Recall that Z-scores are normally distributed with mean 0 and standard deviation 1. Find each of the following probabilities using R code. Note, you should draw a picture to help you visualize what you are trying to find for each question 

    a)	P(Z < 1.25)
    \
    \
    b)	P(Z > 1.25)
    \
    \
    c)	P(Z < -1.25)
    \
    \
    d)	P(Z > -1.25)
    \
    \
    e)	P(-1 < Z < 1)
    \
    \
    f)	P(-2 < Z < 2)
    \
    \
    g)	P(-3 < Z < 3)
    \
    \
    h)	P(-1.645 < Z < 1.645)
    \
    \
    i)	P(-1.96 < Z < 1.96)
    \
    \
    j)	P(-2.576 < Z < 2.576)
\
\
8. Use the qnorm() function to find:

    a)	the 95th percentile of Z-scores
    \
    \
    b)	the 5th percentile of Z-scores
    \
    \
    c)	the 2.5th percentile of Z-scores
    \
    \
    d)	the 97.5th percentile of Z-scores
    \
    \
    e)	the 99.5th percentile of Z-score
    \
    \
9. Suppose that birthweights of newborn babies in the United States follow a normal distribution with mean 3300 grams and standard deviation 500 grams. Babies who weigh less than 2500 grams at birth are classified as low birthweight. Use R code to answer each of the following:

    a)	What percentage of newborn babies weigh less than 2500 grams?
    b)	What is the probability that a randomly selected newborn baby weighs more than 10 pounds?
    c)	What percentage of newborn babies weigh between 3000 and 4000 grams?
    d)	How little must a baby weight to be among the lightest 2.5% of all newborns? Convert your answer to pounds. Google to figure out the conversion!
    e)	How much must a baby weigh to be among the heaviest 10%? Convert your answer to pounds.



\pagebreak

### Gamma example

Suppose $X \sim \Gamma(\alpha = 3, \beta = 4)$.

We can plot this distribution with the following code:

```{r, out.width="50%"}
ggplot() +
  stat_dist_slab(aes(y = "Gamma(3,4)", 
                        #note the function dist_gamma() requires the rate parameter
                        #instead of the scale parameter. Recall rate = 1/scale
                        dist = dist_gamma(shape = 3, rate = 1/4))) +
  labs(y = "") +
  theme_minimal()
```

10.  What's the mean of this distribution? Does this seem reasonable
    based on the graph above?

\
\
\
\

11.  Shade the region $X < 5$ on the graph above

12. Write out the code you need to calculate $P(X < 5)$. *Hint: use the
    `pgamma()` function.* 
    \
    \

13. Use your code above to compute $P(X < 5)$ and report the answer.
    Does the value seem reasonable based on the plot above?\
    \
    \

14. Write out two different lines of code you could use to find
    $P(X > 5)$ in R. *Hint: one should make use of the `lower.tail`
    argument.* Execute both lines of code in R to verify they give the
    value you expect.

\pagebreak

### Chi-squared example

Suppose $X \sim \chi^2(14)$.

We can plot this distribution with the following code:

```{r, out.width="50%"}
ggplot() +
  stat_dist_slab(aes(y = "Chisq(14)", 
                        dist = dist_chisq(14))) +
  labs(y = "", y = "y") +
  theme_minimal()
```

15. What's the mean of this distribution? Does this seem reasonable
    based on the graph above?

\
\

16. Shade the region $X < 10$ on the graph above

17. Write out the code you need to calculate $P(X < 10)$. *Hint: use the `pchisq()` function.* 
\
    \
    \

18. Use your code above to compute $P(X < 10)$ and report the answer.
    Does the value seem reasonable based on the plot above?\
    \

19. Write out two different lines of code you could use to find
    $P(X > 10)$ in R. *Hint: one should make use of the `lower.tail`
    argument.* Execute both lines of code in R to verify they give the
    value you expect.

\pagebreak

### Exponential example

Suppose $X \sim exp(7)$.

We can plot this distribution with the following code:

```{r, out.width="50%"}
ggplot() +
  stat_dist_slab(aes(y = "exp(7)", 
                        #note the function dist_exponential() requires the rate parameter
                        #instead of the scale parameter. Recall rate = 1/scale
                        dist = dist_exponential(1/7))) +
  labs(y = "", y = "y") +
  theme_minimal()
```

20. What's the mean of this distribution? Does this seem reasonable
    based on the graph above?

\
\

21. Write out the code you need to calculate $P(X > 20)$. *Hint: use the
    `pexp()` function.* How many/what arguments do you need to specify?

\
\
\
\

22. Use your code above to compute $P(X > 20)$ and report the answer.
    Does the value seem reasonable based on the plot above?\
    \

23. Shade $P(10 < X < 20)$ on the plot above.

24. Write out the code you need to find $P(10 < X < 20)$. Execute the
    code and report the answer. Does it seem reasonable based on the
    plot above?

\pagebreak

## Computing percentiles for continuous distributions in R

The inverse of the cdf, or the $x$ that solves the equation
$p = F(x) = P(X < x)$, are returned with commands using “q” at the
beginning of the R distribution name:

| Distribution | Quantile Function       | Connection to textbook parametrization |
|---------|-----------------|---------------| 
| Normal       | `qnorm(p, mean, sd)`       | $\mu, \sigma$
| Uniform      | `qunif(p, min, max)`       | $\theta_1, \theta_2$
| Exponential  | `qexp(p, rate)`            | `rate` = $\lambda = 1/\beta$
| Gamma        | `qgamma(p, shape, rate)`  | `shape` = $\alpha$, `rate` = $\lambda = 1/\beta$
| Chi-square   | `qchisq(p, df)`            | $\nu$
| Beta         | `qbeta(p, shape1, shape2)` | $\alpha, \beta$



Again, set `lower.tail = FALSE` to use upper tail probabilities and find
$x$ such that $P(X> x) = p.$

### Gamma example

Suppose $X \sim \Gamma(\alpha = 3, \beta = 7)$. Below is a visual of
the distribution shaded at the median.

```{r, echo = FALSE, out.width="50%"}
median <- qgamma(0.5, shape = 3, scale = 7)
ggplot() +
  stat_dist_slab(aes(y = "gamma(3, 7)", 
                        dist = dist_gamma(shape = 3, rate = 1/7),
                        fill = after_stat(x > median))) +
  scale_fill_viridis_d() +
  labs(y = "") +
  theme_minimal()
```

25. Write out the code you need to find the median in R. Execute the
    code and report the value. Does the value seem reasonable based on
    the plot above?
    
\
\
\
\
26. Write out the code you need to find the 95th percentile in R. Execute the
    code and report the value. Does the value seem reasonable based on
    the plot above?




