---
title: "R discrete distribution cheat sheet"
output: pdf_document
---

# Overview

As you'll recall from Lab 04, R includes many built in functions for common distributions. Commands for each follow the same basic format. The abbreviated name of the distribution is preceded by once of four letters:

-   "d" for the density function (probability distribution for discrete random variables)
-   "p" for the cumulative distribution function (cdf)
-   "q" for the quantile function (gives the x value for a given cumulative probability)
-   "r" for random (generates a random number based on the distribution)

## Binomial distribution

The **binomial distribution** probability distribution and cdf in R:

-   `dbinom(x, size, prob)` -- returns $p(x) = P(X = x),$ (probability distribution) values for a binomial random variable $X$,
-   `pbinom(x, size, prob)` -- returns $F(x) = P(X \leq x),$ values for the binomial distribution (cdf)

Note that both have three arguments defined as:

-   `x` = the value of the random variable for which a probability is computed
-   `size` = number of trials (this is what we usually refer to as $n$)
-   `prob` = probability of success on a trial (what we usually refer to as $p$)

Note that `x` can be a vector, meaning you can compute multiple probabilities with a single command!

## Geometric distribution

**NOTE: In R, the geometric distribution is parameterized as the number of FAILURES until the 1st success.** Whereas, in the book and lecture notes we consider the random variable Y to be the number of the trial on which the first success occurs. This means generally you will need to convert Y to the number of failures, X, as X = Y - 1 before using the built-in R functions.

For example, if you want to find $P(Y = 3)$ (1st success happens on 3rd trial), this is equivalent to $P(X = 2)$ (2 failures before 1st success). This is calculated below, assuming a probability of success of 0.75

```{r}
dgeom(2, 0.75)
```

We can confirm it by calculating in manually:

```{r}
#P(Y = 3) = p(3) = (1 - p)^(3 - 1)*(p)
(1 - .75)^2*.75
```

`pgeom(x, prob)` -- returns $F(x) = P(X \leq x),$ cumulative probabilities for the geometric distribution, again where X = number of failures before 1st success.

$P(Y \leq 3) = P(X \leq 2)$

```{r}
pgeom(2, .75)

#check manually
#P(Y <= 3) = p(1) + p(2) + p(3) = sum((1 - p)^(y - 1)*(p)) for y = 1,2,3
y <- c(1,2,3)
p <- 0.75
p_y <- (1 - p)^(y-1)*p
p_y
sum(p_y)
```

