---
title: "Chapter 3 Part 2 Group Work"
subtitle: "SOLUTIONS"
output: pdf_document
---

## Problem 1

Some biology students were checking eye color in a large number of fruit flies. For the individual fly, suppose that the probability of white eyes is 1/4 and the probability of red eyes is 3/4, and that we may treat these observations as independent Bernoulli trials. What is the probability that at least four flies have to be checked for eye color to observe a white-eyed fly? 

### Solution

$Y \sim geom(p = 0.25)$, where $Y$ is the number of the trial where the first success occurs. 

$P(Y > 3) = (1 - p)^3 = (.75)^3 = 0.421875$

Alternatively, $P(Y \geq 4) = 1 - P(Y < 4) = 1 - p(1) - p(2) - p(3) = 1 - .25 - .75*.25 - .75^2*.25 = 0.421875$

## Problem 2

Suppose that $Y$ is a random variable with a geometric distribution. Show that 

a. $\sum p(y) = \sum_{y = 1}^{\infty}q^{y - 1}p = 1$
b. $\frac{p(y)}{p(y-1)} = q,$ for $y = 2, 3, ....$. This ratio is less than 1, implying that the geometric probabilities are monotonically decreasing as a function of $y$. If $Y$ has a geometric distribution, what value of $Y$ is the most likely (has the highest probability)?

### Solution

#### Part a

$$
\begin{aligned}
\sum_{y = 1}^{\infty}pq^{y - 1} = p\sum_{y - 1 = 0}^{\infty}q^{y - 1} = p\frac{1}{1 - q} = \frac{p}{1 - (1-p)} = \frac{p}{p} = 1
\end{aligned}
$$

#### Part b

Since $p(y)$ is monotonically decreasing, and the support of $Y$ is $1, 2, 3, \dots$, then the largest probability is when $Y = 1$. 

## Problem 3

About 7 months into Donald Trump's 2nd term as president (August 2025), a Gallup poll found that a record low of 39% of adults approved of how the Supreme Court is handling its job. 

a. Find the probability distribution for $Y$, the number of calls until the first person is found who *does* express approval of the U.S. Supreme Court. 
b. On average, how many calls are needed until the 1st approval is found?
c. Find the probability distribution for $Z$, the number of calls until the 50th person is found who approves of the U.S. Supreme Court. 
d. On average, how many calls are needed until the 50th approval is found? 

### Solution

#### Part a

$Y\sim geom(p = 0.39) \implies p(y) = (.61)^{y - 1}(.39)$

#### Part b

$E(Y) = \frac{1}{p} = \frac{1}{.39} = 2.56$

#### Part c

$Z \sim nbinom(0.39, 50) \implies p(z)= \binom{z - 1}{49}(.39)^{50}(.61)^{z - 50}$

#### Part d

$E(Z) = \frac{r}{p} = \frac{50}{.39} = 128.2$

## Problem 4

The employees of a firm that manufactures insulation are being tested for indications of asbestos in their lungs. The firm is requested to send three employees who have positive indications of asbestos on to a medical center for further testing. If 40% of the employees have positive indications of asbestos in their lungs, find the probability that 10 employees must be tested in order to find three positives. 

### Solution

$Y \sim nbinom(r = 3, p = 0.4)$

$P(Y = 10) = \binom{9}{2}(0.4)^3(0.6)^7 = 0.06449$

## Problem 5

A jury of 6 persons was selected from a group of 20 potential jurors, of whom 8 were Black and 12 were White. The jury was supposedly randomly selected, but it contained only 1 Black member. Do you have any reason to doubt the randomness of the selection? 

### Solution

$$
\frac{\binom{8}{1}\binom{12}{5}}{\binom{20}{6}} = \frac{8*792}{38760} = 0.163
$$

There is about a 16% chance of this happening, which is not particularly rare. There is not strong evidence to doubt the randomness of selection in this case.

## Problem 6

The number of typing errors made by a typist has a Poisson distribution with an average of four errors per page. If more than four errors appear on a given page, the typist must retype the whole page. What is the probability that a randomly selected page does not need to be retyped?

### Solution

Unit of interest = 1 page. $Y \sim Poisson(\lambda = 4)$.

$P(Y \leq 4) = \sum_{y = 0}^4 \frac{4^ye^{-4}}{y!} = e^{-4}[4^0/0! + 4/1! + 4^2/2! + 4^3/3! + 4^4/4!] = 0.6288$

An example of streamlining this calculation in R:

```{r}
y <- 0:4
lambda <- 4
p_y <- exp(-4)*4^y/factorial(y)
p_y
sum(p_y)
```


## Problem 7

Let $Y$ be a random variable with probability distribution $p(y) = \frac{y}{6}, \ \ \ y = 1, 2, 3$

  a) Find an expression for the moment generating function of $Y$. That is, write $E(e^{tY})$ as a sum.
  b) Use the mgf to show that $E(Y) = 7/3$
  c) Use the mgf to show that $E(Y^2) = 6$
  d) Find $V(Y)$

### Solution

#### part a 

$$\begin{aligned} 
E(e^{tY}) &= \sum e^{ty}f(y) \\ 
&= \sum e^{ty}\left(\frac{y}{6}\right) 
\end{aligned}$$ 

#### part b 

Recall that $E(Y)$ is the "first moment." So we need to find the first derivative of the mgf and evaluate it at $t = 0$ 

$$\begin{aligned} 
m(t) &=\sum_{y = 1}^3 e^{ty}\left(\frac{y}{6}\right) \\ 
m'(t) &= \sum_{y = 1}^3 e^{ty}(y)\left(\frac{y}{6}\right) \\ 
E(y) =m'(0) &= \sum_{y = 1}^3 e^{0y}(y)\left(\frac{y}{6}\right) \\ 
&= \sum_{y = 1}^3 \frac{y^2}{6} \\ 
&= \frac{1}{6}\sum_{y = 1}^3 y^2 \\ 
&= \frac{1}{6}[1^2 + 2^2 + 3^2] \\ 
&= \frac{14}{6} \\ 
&= \frac{7}{3} 
\end{aligned}$$ 

#### part c 

Recall that $E(y^2)$ is the "second moment" so we need to take the second derivative of $m(t)$ and evaluate it at 0. 
$$\begin{aligned} 
m'(t) &= \sum_{y = 1}^3 e^{ty}\left(\frac{y^2}{6}\right) \\ 
m''(t) &= \sum_{y = 1}^3 e^{ty}(y)\left(\frac{y^2}{6}\right) \\ 
m''(0) &= \sum_{y = 1}^3 e^{0y}(y)\left(\frac{y^2}{6}\right) \\ 
E(y^2) = m''(0) &= \sum_{y = 1}^3 \frac{y^3}{6} \\ 
&= \frac{1^3}{6} + \frac{2^3}{6} + \frac{3^3}{6} \\ 
&= 6 
\end{aligned}$$ 

#### part d 

$$V(y) = E(y^2) - [E(y)]^2 = 6 - \left(\frac{7}{3}\right)^2 = \frac{5}{9}$$ 
  
## Problem 8

Obtain an expression for the mgf of the Poisson distribution. Use the mgf to show that $E(Y) = V(Y) = \lambda$ for a Poisson random variable.

### Solution

Let $Y$ be Poisson with parameter $\lambda>0$. The probability distribution is

$$
p(y) = \frac{e^{-\lambda}\lambda^{y}}{y!}, \qquad y=0,1,2,\dots
$$

The moment generating function is defined as 

$$
\begin{aligned}
m(t) &= E\big(e^{tY}\big) = \sum_{y=0}^{\infty} e^{ty}\ p(y) \\
&= \sum_{y=0}^{\infty} e^{ty}\,\frac{e^{-\lambda}\lambda^{y}}{y!} \\[6pt]
&= e^{-\lambda}\sum_{y=0}^{\infty}\frac{(\lambda e^{t})^{y}}{y!} \\[6pt]
&= e^{-\lambda}\exp(\lambda e^{t}) \\[6pt]
&= \exp\!\big(\lambda(e^{t}-1)\big).
\end{aligned}
$$

Take the first derivative to find the mean

$$
M_Y'(t) = \lambda e^{t}\exp\!\big(\lambda(e^{t}-1)\big).
$$

Evauluate at $t=0$:

$$
E(Y) = M_Y'(0) = \lambda.
$$

Take the second derivative to find $E(Y^2$):

$$
m''(t) = \lambda e^{t}\exp\!\big(\lambda(e^{t}-1)\big)\big(1+\lambda e^{t}\big).
$$

Evaluate at $t=0$:

$$
M_Y''(0) = \lambda(1+\lambda).
$$

Thus,

$$
\operatorname{Var}(Y) = m''(0) - \big(m'(0)\big)^2
= \lambda(1+\lambda) - \lambda^2
= \lambda.
$$

## Problem 7

In a lab experiment involving inorganic syntheses of molecular precursors to organometallic ceramics, the final step of a five-step reaction involves the formation of a metal-metal bond. The probability of such a bond forming is $p = 0.2$. Let $X$ equal the number of successful reactions out of $n=25$ such experiments.

a) What distribution is appropriate to model $X$?
b) Write out the pmf of $X$
c) Find the probability that $X = 1$
d) Find the probability that $X$ is at least 1
e) Find the mean, variance, and standard deviation of $X$

### Solution part a

$$X \sim binomial(25, 0.2)$$

### Solution part b

$$p(x) = {25 \choose x}(0.2)^x(0.8)^{n-x}$$

### Solution part c

$$\begin{aligned} P(X =1) = f(1) &= {25 \choose x}(0.2)^x(0.8)^{25-x} \\
&= {25 \choose 1}(0.2)^1(0.8)^{25-1} \\
&= 25(.2)(.8)^{24} \\
&= 0.0236
\end{aligned}$$

### Solution part d

$$\begin{aligned}
P(X \geq 1) &= 1- P(X = 0) \\
&= 1 - {25 \choose 0}(0.2)^0(0.8)^{25-0} \\
&= 1 - 1(1)(.8)^{25} \\
&= 0.9962
\end{aligned}$$

### Solution part e

Recall that $E(X) = np$ and $V(X) np(1-p)$ for a binomial distribution. Therefore in this case, we have

$$E(X) = 25(0.2)=5$$
$$\sigma^2 = V(X) = 25(0.2)(0.8)=4$$

$$\sigma = \sqrt{V(X)} = \sqrt{4} = 2$$

##  Problem 8

A random variable X has a binomial distribution with mean 6 and variance 3.6. Find $P(X = 4)$.

### Solution

We are given that $E(X) = np = 6$ and $V(X) = np(1-p) = 3.6$. Therefore, we can plug in and set $6(1-p) = 3.6$ Solving for $p$ gives $p = 0.4$. Plugging back into $np = 6$ gives $n = 15$. We can then write the pmf of $X$ as

$$f(x) = {15 \choose x} p^x(1-p)^{15-x}$$
$$P(X = 4) = f(4) = {15 \choose 4} (0.4)^4(0.6)^{11} = 0.1268$$


## Problem 10

Return to the Joel Embiid example (he makes 87% of his free throws, he's fouled shooting a 3-pointer). Describe in detail how you could, in principle, perform a by hand simulation involving physical objects (e.g. coins, dice, spinners, etc.) to estimate the probability that the Lakers lose. Be sure to describe (1) what one repetition of the simulation entails, and (2) how you would use the results of many repetitions. Note: you do NOT need to compute any numerical values.

### Solution

[Answers may vary]. One possible solution:

(1) To simulate one repetition of the experiment, we could use a spinner where 87% of the area is shaded as "success" and 13% of the area is shaded as "failure". We spin the spinner 3 times and record the number of successes. This gives us one value of the random variable $X$

(2) We could repeat the process in part (1) 100,000 times. This would give us 100,000 values of the random variable $X$. Note that the Sixers lose if $X = 0$ or $X = 1$, they tie if $X =2$, and they win if $X = 3$. So to find the probability that they lose, we count up the number of times $X$ equals 0 or 1 and divide by 100,000.

### Additional fun

The `rbinom()` function in R simulates values from the binomial distribution, so we could run the above simulations (and save them into a data frame called `data`) with the following code:

```{r}
data <- data.frame(x = rbinom(100000, 3, 0.87))
```

A bar plot of the 100,000 simulated $x$ values shows us an approximate distribution of $X$

```{r, out.width="60%"}
ggplot(data = data, aes(x = x)) +
  geom_bar() +
  theme_classic()
```

To estimate the probability the Lakers lose from our simulations, we have:

```{r}
data %>%
  summarize(prop = sum(x == 0 | x == 1)/100000)
```

We can compare this to the theoretical probability from the pmf:

```{r}
pbinom(1, 3, .87)
```

Here is the true pmf (navy outline) plotted on top of the simulated distribution (grey fill), and we can see they are nearly identical.

```{r}
#convert simulated counts into proportions & compute pmf for each value of x
data_summary <- data %>%
  count(x, name = "count") %>%
  mutate(rel_freq = count/100000,
         pmf = dbinom(x, 3, .87))

ggplot(data_summary) +
  geom_col(aes(x = factor(x), y = rel_freq),
           alpha = 0.8) +
  geom_col(aes(x = factor(x), y = pmf),
           fill = NA, color = "navyblue") +
  labs(x = "x", y = "Density f(x)") +
  theme_classic()
```





