---
format: pdf
fontsize: 11pt
geometry: margin=0.5in
---
## 1. R Functions for the Binomial Distribution

R uses a consistent naming scheme for distributions:

- d: density / pmf → `dbinom()`
- p: cumulative probabilities / cdf → `pbinom()`
- q: quantiles → `qbinom()`
- r: random generation → `rbinom()`

For Binomial:

- `dbinom(x, size, prob)` gives $P(X = x)$
- `pbinom(q, size, prob)` gives $P(X ≤ q)$

Arguments:

- `x,q` = value(s) of the random variable, number of successes
- `size` = number of trials n
- `prob` = success probability p

Note: `x` or `q` may be a single value or a vector to compute multiple probabilities at once.

## 2. Quick Examples (with X ~ Binomial(5, 0.05))

No defective items ($P(X = 0)$)
```{r}
dbinom(x = 0, size = 5, prob = .05)
```


1 or fewer defective items ($P(X \leq 1)$)
```{r}
pbinom(q = 1, size = 5, prob = .05)
```


```{r}
x <- 0:5
dbinom(x, 5, 0.05) #All pmf values for x = 0–5
```


```{r}
sum(dbinom(x, 5, 0.05)) #Check pmf sums to 1
```



```{r}
dbinom(2, 5, 0.05) #Exactly 2 successes
```


```{r}
pbinom(2, 5, 0.05) #At most 2 successes
```


```{r}
1 - pbinom(2, 5, 0.05) #More than 2 successes
pbinom(2, 5, .05, lower.tail = FALSE) #Alternative way to find more than 2 successes:
```

```{r}
library(ggplot2)

data <- data.frame(x = x, d = dbinom(x, 5, 0.05))

ggplot(data) +
  geom_col(aes(x = factor(x), y = d), fill = "skyblue") +
  labs(x = "x", y = "Density f(x)") +
  theme_classic()
```

<!-- --- -->

<!-- ## 3. Binomial Mean & Variance -->

<!-- If X ~ Binomial(n, p): -->

<!-- - Mean: μ = np -->
<!-- - Variance: σ² = np(1−p) -->
<!-- - Standard deviation: σ = sqrt(np(1−p)) -->

<!-- --- -->

<!-- ## 4. Example (Use as a Template) -->

<!-- A mint weighs more than 20.7g with probability 0.90. Let X = number of heavy mints in a sample of 8. Then X ~ Binomial(8, 0.9). -->

<!-- a. Mean & SD -->
<!-- μ = 8(0.9) -->
<!-- σ = sqrt(8(0.9)(0.1)) -->

<!-- ```{r} -->
<!-- mu <- 8*0.9 -->
<!-- sigma <- sqrt(8*0.9*0.1) -->
<!-- mu; sigma -->
<!-- ``` -->

<!-- b. P(X = 8) -->
<!-- ```{r} -->
<!-- dbinom(8, 8, 0.9) -->
<!-- ``` -->

<!-- c. P(X ≤ 6) -->
<!-- ```{r} -->
<!-- pbinom(6, 8, 0.9) -->
<!-- ``` -->

<!-- --- -->

<!-- ## 5. Practice Problems -->

<!-- ### Task 1 -->
<!-- X ~ Binomial(25, 0.20) -->

<!-- Compute: -->
<!-- - P(X ≤ 4) -->
<!-- - P(X ≥ 5) -->
<!-- - P(X = 6) -->
<!-- - Mean, variance, standard deviation -->

<!-- ```{r} -->
<!-- p_at_most_4 <- pbinom(4, 25, 0.20) -->
<!-- p_at_least_5 <- 1 - pbinom(4, 25, 0.20) -->
<!-- p_eq_6 <- dbinom(6, 25, 0.20) -->

<!-- mu_1  <- 25*0.20 -->
<!-- var_1 <- 25*0.20*0.80 -->
<!-- sd_1  <- sqrt(var_1) -->

<!-- p_at_most_4; p_at_least_5; p_eq_6 -->
<!-- mu_1; var_1; sd_1 -->
<!-- ``` -->

<!-- --- -->

<!-- ### Task 2 -->
<!-- X ~ Binomial(15, 0.75) -->

<!-- Compute: -->
<!-- - P(X ≥ 10) -->
<!-- - P(X ≤ 10) -->
<!-- - P(X = 10) -->
<!-- - Mean, variance, standard deviation -->

<!-- ```{r} -->
<!-- p_at_least_10 <- 1 - pbinom(9, 15, 0.75) -->
<!-- p_at_most_10  <- pbinom(10, 15, 0.75) -->
<!-- p_eq_10       <- dbinom(10, 15, 0.75) -->

<!-- mu_2  <- 15*0.75 -->
<!-- var_2 <- 15*0.75*0.25 -->
<!-- sd_2  <- sqrt(var_2) -->

<!-- p_at_least_10; p_at_most_10; p_eq_10 -->
<!-- mu_2; var_2; sd_2 -->
<!-- ``` -->